---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Install Libraries
```{r, message=FALSE}
library(dplyr)
library(ggplot2)
```
## Loading and preprocessing the data
Download and read files
```{r read}
activity <- read.csv(unz("activity.zip", "activity.csv"), colClasses = c("integer","character","integer"))
activity$date <- as.Date(activity$date, "%Y-%m-%d")
activity <- tbl_df(activity)
```

## What is mean total number of steps taken per day?
Summarize the activity table by total steps per day.
```{r}
day <- group_by(activity,date)
daytotals <- summarize(day,steps = sum(steps, na.rm = TRUE))
```
Calculate mean and median.
```{r}
mean <- mean(daytotals$steps)
median <- median(daytotals$steps)
```
Plot histogram with a vertical line showing the median and mean.
```{r histogram, fig.height=6,fig.width=10}
ggplot(daytotals, aes(x = steps)) +
        geom_histogram(bins = 30) +
        geom_vline(xintercept = mean, color = "red") +
        geom_text(aes(
                x = mean - 2000,
                y = 10,
                label = paste("mean =", round(mean, 0))
                
        ), color = "red") +
        geom_text(aes(
                x = median + 2000,
                y = 10,
                label = paste("median =", round(median, 0))
                
        ), color = "blue") +
        geom_vline(xintercept = median, color = "blue") +
        xlab("Total Daily Steps") +
        ylab("Number of Days") +
        ggtitle("Distribution of Total Daily Steps") +
        theme(plot.title = element_text(hjust = 0.5))
```

The mean and median number of steps per day are `r round(mean,0)` and `r median` respectively.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
